---
title: "Kenya EIE Working Group -- UNICEF sitrep inputs"
date:  "31 December 2023"
output: 
  word_document:
    toc: no
always_allow_html: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=9, message = FALSE, warning=FALSE)
library(tidyverse)
library(readxl)
library(lubridate)
library(stringi)
library(janitor)
library(scales)
library(viridis)
library(patchwork)
library(DT)
library(here)
library(sf)
library(plotly)
library(flextable)
library(ggrepel)
library(anytime)
library(tidytext)

theme_set(theme_light())

# Installation of mdepriv package
# install package devtools if not yet installed
#   install.packages("devtools")

# install fast from GitHub without vignettes 
#   devtools::install_github("a-benini/mdepriv", build_vignettes = FALSE

# Alternatively, follow the instructions here: https://a-benini.github.io/mdepriv/

# disabling scientific notation
options(scipen = 100)

`%out%` <- Negate(`%in%`)

# function for transposing df
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
  return(t_df)
}

# scaling functions 
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
range_wna <- function(x){(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))}

# mode function 
mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Fixing dates
name_to_date <- function(x) {
  lubridate::mdy(ifelse(x %in% c(month.name, month.abb), paste0(x, "/01/2023"), x))
}

# Recode last item
# Doesn't work well
recode_last_item <- function(input_vector) {
  # Create a copy of the input vector
  updated_vector <- input_vector
  
  new_value <- paste0("and ", updated_vector[length(updated_vector)])
  
  # Recode the last item
  updated_vector[length(updated_vector)] <- new_value 
  
  # Return the updated vector
  return(updated_vector)
  
}


 summarise_indicator <- function(tbl) {
  tbl %>% 
    summarise(girls = sum(girls, na.rm = TRUE), 
            boys = sum(boys, na.rm = TRUE), 
            men = sum(men, na.rm = TRUE), 
            women = sum(women, na.rm = TRUE), 
            counties = n_distinct(adm1_pcode)) %>%
    mutate_at(vars(girls, boys, men, women), ~ replace_na(., 0)) %>% 
    mutate(total = girls + boys + men + women)
}

list_counties <- function(tbl) {
  tbl %>% 
    distinct(county) %>% summarise(
    counties = paste(county, collapse = ", ")) %>%
    pull(counties)
}


```




```{r}
locations <- read_xlsx("./data/ken_adminboundaries_tabulardata.xlsx", 
                      sheet = "ADM2") %>% 
  clean_names()

eie <- read_csv("./data/eie_5ws.csv")

report_date <- "31 December 2023"

previous_months <- c("January", "February", "March", "April", "May", 
                     "July", "August", "September", "October", "November")

report_month <- "December"

sitrep_table <- read_xlsx("./data/sitrep_results_table.xlsx") %>% 
  setNames(c(
    "indicator", 
    "indicator_match", 
    "sex_modifier", 
    "needs", 
    "unicef_target", 
    "unicef_total_results", 
    "unicef_previous_results", 
    "unicef_progress_rate", 
    "sector_target", 
    "sector_total_results", 
    "sector_previous_results", 
    "sector_progress_rate"
  ))

# Check the file name
# combined <- read_csv("./data/combined20230904.csv")
```



# Sitrep 



```{r}
indicator1 <- eie %>% 
  replace_na(list(boys = 0, girls = 0, men = 0, women = 0)) %>% 
  # Total reached of indicator 1 is only boys and girls 
  mutate(total_reached = ifelse(str_detect(indicator, "1"), 
                                boys + girls, 
                                total_reached)) %>% 
  filter(month %in% report_month & str_detect(indicator, "1") & 
                 lead_organisation == "UNICEF" & activity_status == "Completed") %>% 
  summarise(girls = sum(girls, na.rm = TRUE), 
          boys = sum(boys, na.rm = TRUE), 
          men = sum(men, na.rm = TRUE), 
          women = sum(women, na.rm = TRUE), 
          counties = n_distinct(adm1_pcode)) %>%
  mutate_at(vars(girls, boys, men, women), ~ replace_na(., 0)) %>% 
  mutate(total = girls + boys)

indicator1_counties <- eie %>% filter(month %in% report_month & str_detect(indicator, "1") & 
                 lead_organisation == "UNICEF" & activity_status == "Completed") %>% 
  list_counties() 

```

A total of `r indicator1 %>% pull(total) %>% format(big.mark = ",")` children (`r indicator1 %>% pull(girls) %>% format(big.mark = ",")` girls and `r indicator1 %>% pull(boys) %>% format(big.mark = ",")` boys) have accessed basic education with UNICEF support during `r report_month` 2023 in the counties of `r indicator1_counties`. 


```{r}
indicator3 <- eie %>% 
  replace_na(list(boys = 0, girls = 0, men = 0, women = 0)) %>% 
  # Total reached of indicator 1 is only boys and girls 
  mutate(total_reached = ifelse(str_detect(indicator, "3"), 
                                boys + girls, 
                                total_reached)) %>% 
  filter(month %in% report_month & str_detect(indicator, "3") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>% 
  summarise(girls = sum(girls, na.rm = TRUE), 
          boys = sum(boys, na.rm = TRUE), 
          men = sum(men, na.rm = TRUE), 
          women = sum(women, na.rm = TRUE), 
          counties = n_distinct(adm1_pcode)) %>%
  mutate_at(vars(girls, boys, men, women), ~ replace_na(., 0)) %>% 
  mutate(total = girls + boys)


indicator3_counties <- eie %>% 
  filter(month %in% report_month & str_detect(indicator, "3") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>%
  list_counties()
```



Furthermore, `r indicator3 %>% pull(total) %>% format(big.mark = ",")` students (`r indicator3 %>% pull(girls) %>% format(big.mark = ",")` girls and `r indicator3 %>% pull(boys) %>% format(big.mark = ",")` boys) in `r indicator3_counties` County benefited from distribution of school bags and education kits to support continuation of their learning. 
 

```{r}
indicator5.1 <- eie %>% 
  replace_na(list(boys = 0, girls = 0, men = 0, women = 0)) %>% 
  # Total reached of indicator 1 is only boys and girls 
  mutate(total_reached = ifelse(str_detect(indicator, "5.1"), 
                                men + women, 
                                total_reached)) %>% 
  filter(month %in% report_month & str_detect(indicator, "5.1") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>% 
  summarise(girls = sum(girls, na.rm = TRUE), 
          boys = sum(boys, na.rm = TRUE), 
          men = sum(men, na.rm = TRUE), 
          women = sum(women, na.rm = TRUE), 
          counties = n_distinct(adm1_pcode)) %>%
  mutate_at(vars(girls, boys, men, women), ~ replace_na(., 0)) %>% 
  mutate(total = men + women)

```
 

To enhance quality of learning and support newly enrolled learners in drought affected counties, `r indicator5.1 %>% pull(total) %>% format(big.mark = ",")` teachers (`r indicator5.1 %>% pull(men) %>% format(big.mark = ",")` men and `r indicator5.1 %>% pull(women) %>% format(big.mark = ",")` women) were trained on remedial and catchup pedagogical approaches, Mental Health and Psychosocial Support (MHPSS), Gender-Based Violence (GBV) and Guidance and Counselling (G&C). 


```{r}
indicator5.2 <- eie %>% 
  replace_na(list(boys = 0, girls = 0, men = 0, women = 0)) %>% 
  # Total reached of indicator 1 is only boys and girls 
  mutate(total_reached = ifelse(str_detect(indicator, "5.2"), 
                                men + women, 
                                total_reached)) %>% 
  filter(month %in% report_month & str_detect(indicator, "5.2") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>% 
  summarise(girls = sum(girls, na.rm = TRUE), 
          boys = sum(boys, na.rm = TRUE), 
          men = sum(men, na.rm = TRUE), 
          women = sum(women, na.rm = TRUE), 
          counties = n_distinct(adm1_pcode)) %>%
  mutate_at(vars(girls, boys, men, women), ~ replace_na(., 0)) %>% 
  mutate(total = men + women)

indicator5.2_counties <- eie %>% 
  filter(month %in% report_month & str_detect(indicator, "5.2") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>%
  list_counties()
  
  
```


UNICEF also supported the training of `r indicator5.2 %>% pull(total) %>% format(big.mark = ",")` (`r indicator5.2 %>% pull(men) %>% format(big.mark = ",")` men and `r indicator5.2 %>% pull(women) %>% format(big.mark = ",")` women) members of the Enrolment Drive Committees (EDCs) and school Board of Management (BoM) in `r indicator5.2_counties` county on life skills, child safeguarding and their roles/responsibilities, child protection and referral mechanisms. 


```{r}
indicator6 <- eie %>% 
  replace_na(list(boys = 0, girls = 0, men = 0, women = 0)) %>% 
  # Total reached of indicator 1 is only boys and girls 
  mutate(total_reached = ifelse(str_detect(indicator, "6"), 
                                boys + girls, 
                                total_reached)) %>% 
  filter(month %in% report_month & str_detect(indicator, "6") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>% 
  summarise(girls = sum(girls, na.rm = TRUE), 
          boys = sum(boys, na.rm = TRUE), 
          men = sum(men, na.rm = TRUE), 
          women = sum(women, na.rm = TRUE), 
          counties = n_distinct(adm1_pcode)) %>%
  mutate_at(vars(girls, boys, men, women), ~ replace_na(., 0)) %>% 
  mutate(total = girls + boys)

indicator6_counties <- eie %>% 
  filter(month %in% report_month & str_detect(indicator, "6") & 
           lead_organisation == "UNICEF" & activity_status == "Completed") %>%
  list_counties()


```
 
 

Furthermore, `r indicator6 %>% pull(total) %>% format(big.mark = ",")` children (`r indicator6 %>% pull(girls) %>% format(big.mark = ",")` girls and `r indicator6 %>% pull(boys) %>% format(big.mark = ",")` boys) in `r indicator6_counties` were supported to access mentorship and life skills sessions in `r indicator6_counties` counties, strengthening their resilience and their coping mechanisms to the impact of drought.

\newpage 

# Sitrep table 



```{r}
sitrep_table_new <- eie %>%
      filter(activity_status == "Completed") %>% 
      mutate(group = ifelse(str_detect(lead_organisation, "UNICEF"), 
                             "unicef", "non_unicef")) %>%
      filter(str_detect(indicator, "1|3") & 
               !str_detect(indicator, "5")) %>%
      mutate(total = boys + girls) %>% 
      pivot_longer(cols = c(boys, girls, total), 
                   names_to = "sex_modifier", 
                   values_to = "beneficiaries") %>% 
      mutate(sex_modifier = str_to_title(sex_modifier), 
             sex_modifier = fct_relevel(sex_modifier, 
                                        c("Total", "Girls", "Boys"))) %>% 
      group_by(indicator_short, sex_modifier) %>%
      summarise(unicef_previous_results = sum(beneficiaries[month %in% previous_months & group == "unicef"], 
                                      na.rm = TRUE),
                unicef_total_results = sum(beneficiaries[group == "unicef"], 
                                      na.rm = TRUE),
                sector_previous_results = sum(beneficiaries[month %in% previous_months], na.rm = TRUE),
                sector_total_results = sum(beneficiaries, na.rm = TRUE), 
                .groups = "drop") %>% 
      mutate(indicator_match = ifelse(indicator_short == "1. Access ECD spaces/schools", 
                                      "Access to education", 
                                      "Learning materials"), 
             match = paste0(indicator_match, sex_modifier)) %>%
  left_join(sitrep_table %>%
              mutate(indicator_match = ifelse(str_detect(indicator_match, "education"), 
                                                         "Access to education",
                                                         "Learning materials"),
                                              sex_modifier = case_when(
                                                str_detect(sex_modifier, "Total") ~ "Total",
                                                str_detect(sex_modifier, "Girls") ~ "Girls",
                                                str_detect(sex_modifier, "Boys") ~ "Boys")) %>%
              # You can remove this later, since you've figured out the error
                       mutate(match = paste0(indicator_match, sex_modifier)) %>%
                       mutate_at(vars(needs, unicef_target, sector_target),
                                                      ~ parse_number(.)) %>%
                       select(match, needs, unicef_target, sector_target), 
            by = c("match")) %>% 
  mutate(unicef_progress_rate = round(unicef_total_results / unicef_target * 100, digits = 2), 
         sector_progress_rate = round(sector_total_results / sector_target * 100, digits = 2)) %>% 
  select(indicator = indicator_match, sex_modifier, needs, 
         unicef_target, unicef_total_results, unicef_previous_results, unicef_progress_rate, 
         sector_target, sector_total_results, sector_previous_results, sector_progress_rate) 

sitrep_table_new%>% 
  flextable() %>% 
  theme_zebra() %>% 
  fontsize(part = "header", size = 6) %>% 
  fontsize(i = 1:6, j = 1:11, size = 7)

sitrep_table_new %>% 
  write_csv(paste0("./data/sitrep_table_new_", report_month, ".csv"))

```

## Downloads

[5ws dataset](https://github.com/kenya-eie-wg/eie_wg_5ws/blob/main/data/eie_5ws.csv)








